NanoMaLy

### 2015-11-15 paper/pldi16-submission/interactive.tex
more interactive

First, we extend the transition relation to collect a set of edges from
which we will construct the graph. 
%
An edge between two expressions is either a ``steps-to'' edge -- written
\singlestep{e_1}{e_2} -- indicating that $e_1$ transitions to $e_2$ in a
single step, or a ``sub-term'' edge -- written \subterm{e_1}{e_2} --
indicating that $e_1$ contains $e_2$ as a sub-expression.
%
Collecting the steps-to edges is a simple matter of recording the
consequent of each original rule in the trace; each original judgment
\step{e_1}{\su_1}{e_2}{\su_2} becomes 
\stepg{e_1}{\su_1}{\tr_1}{e_2}{\su_2}{\singlestep{e_1}{e_2}; \tr_1}.
%
The sub-term edges can be delegated to a new \subtermssym helper
function, which adds edges from an expression to each of its
\emph{immediate} sub-expressions.
%
We collect new sub-term edges after each transition, thus the final
template for the small-step relation is:
$$
\stepg{e_1}{\su_1}{\tr_1}{e_2}{\su_2}{\singlestep{e_1}{e_2}; \subterms{e_2}; \tr_1}
$$

% \begin{itemize}
% \item extend operational semantics to collect reduction graph
% \item nodes are terms, edges indicate ``steps-to'' and ``sub-term'' relations
% \item visualize path through reduction graph
% \item expand edges to reveal more fine-grained steps (step/jump forward/backward)
% \item never lose context (unlike traditional debugger)
% \end{itemize}

>>>>

First, we extend the transition relation to collect a set of edges from
which we will construct the graph.
%
An edge between two expressions is either a ``steps-to'' edge -- written
\singlestep{e_1}{e_2} -- indicating that $e_1$ transitions to $e_2$ in a
single step, or a ``sub-term'' edge -- written \subterm{e_1}{e_2} --
indicating that $e_1$ contains $e_2$ as a sub-expression.
%
Collecting the steps-to edges is a simple matter of recording the
consequent of each original rule in the trace; each original judgment
\step{e_1}{\su_1}{e_2}{\su_2} becomes
\stepg{e_1}{\su_1}{\tr_1}{e_2}{\su_2}{\singlestep{e_1}{e_2}; \tr_1}.
%
The sub-term edges can be delegated to a new \subtermssym helper
function, which adds edges from an expression to each of its
\emph{immediate} sub-expressions.
%
We collect new sub-term edges after each transition, thus the final
template for the small-step relation is:
\[
\stepg{e_1}{\su_1}{\tr_1}{e_2}{\su_2}{\singlestep{e_1}{e_2}; \subterms{e_2}; \tr_1}
\]

\subsection{Traversing the Reduction Graph}
\label{sec:traversing-graph}
Given a reduction graph $G$ and an initial path
$p = \singlestep{e_1}{\singlestep{e_2}{e_n}}$, such that
\stepsg{e_1}{\emptysu}{\bullet}{e_n}{\su}{\tr}, we define the following
traversals in Figure~\ref{fig:traversing-graph}:
%
\begin{itemize}
\item \stepforwardsym takes a single step forward.
\item \stepbackwardsym takes a single step backward.
\item \jumpforwardsym takes a ``big'' step forward to the next function call.
\item \jumpbackwardsym takes a ``big'' step backward to the previous function call.
\item \stepintosym steps into a function call in a sub-term, isolating it from the context.
\item \stepoversym steps over a function call in a sub-term.
\end{itemize}
%
The initial path $p$ is required for the backward-steps as a node may
have multiple incoming $\leadsto$ edges, \eg
\singlestep{\eplus{1}{2}}{3} and \singlestep{\eplus{2}{1}}{3}.
%
The sub-term edges $\searrow$ allow us to decompose an expression into a
sub-expression and the surrounding context, thus enabling the \stepintosym
and \stepoversym traversals.

\begin{figure*}[t]
\centering
\[
\begin{array}{lcl}
\stepforward{e_i}  &\defeq& {e_{i+1}} \\ \\
\stepbackward{e_i} &\defeq& {e_{i-1}} \\ \\
\jumpforward{e_i} &\defeq& \left\{\begin{array}{ll}
                         e_{i+1}, & \text{if } e_{i+1} = \eapp{v_1}{v_2} \\
                         \jumpforward{e_{i+1}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\jumpbackward{e_i} &\defeq& \left\{\begin{array}{ll}
                         e_{i-1}, & \text{if } e_{i-1} = \eapp{v_1}{v_2} \\
                         \jumpbackward{e_{i-1}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\stepinto{e_i} &\defeq& \left\{\begin{array}{ll}
                         e\sub{x}{v_2}, & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}
                         \end{array}\right\} \\ \\
\stepover{e_i} &\defeq& \left\{\begin{array}{ll}
                         C[v], & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \multistep{\eapp{v_1}{v_2}}{v}
                         \end{array}\right\}
\end{array}
\]
\caption{Rules for traversing the reduction graph. \stepintosym and \stepoversym
  require a traversal of the sub-term edges to decompose $e_i$ into the
  target expression \eapp{v_1}{v_2} and the context $C$.}
\label{fig:traversing-graph}
\end{figure*}
% \begin{itemize}
% \item extend operational semantics to collect reduction graph
% \item nodes are terms, edges indicate ``steps-to'' and ``sub-term'' relations
% \item visualize path through reduction graph
% \item expand edges to reveal more fine-grained steps (step/jump forward/backward)
% \item never lose context (unlike traditional debugger)
% \end{itemize}
